\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2018-09-27 redaction/online_appendix_travail.tex
fin modifs OA par JH

    Std & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline

@@

    Std & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.01} \\ \hline

@@

    Std & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.10} & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.06} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.05} \\  \hline \hline

@@

\begin{table}[htbp]
  \centering
   \footnotesize{
    \caption{Estimation results of the three models (Vessel, 3-digit level)}
  \label{tab:3models_estimation_results_vessel}%
    \begin{tabular}{l|c c c c c c c c c}
    \hline \hline
    \textbf{Year} & 1974  & 1980  & 1985  & 1990  & 1995  & 2000  & 2005  & 2010  & 2013 \\ \hline
    $\# $obs. & 19007 & 17356 & 23348 & 28383 & 32146 & 36090 & 41319 & 37748 & 38473 \\
    $\#$  origin countries & 154   & 163   & 171   & 179   & 201   & 206   & 206   & 198   & 203 \\
    $\#$  sectors & 239   & 232   & 232   & 232   & 228   & 230   & 231   & 226   & 224 \\ \hline
    \multicolumn{10}{l}{\textbf{Model (A) - Iceberg transport costs only} ($\widehat{\tau}^{ice}$)} \\ \hline
    Mean (in \%) & 9.79  & 6.53  & 6.88  & 5.67  & 5.14  & 5.10  & 5.47  & 3.99  & 3.60 \\
    Median (in \%)& 9.58  & 5.50  & 6.33  & 4.63  & 4.29  & 4.85  & 4.90  & 3.56  & 3.28 \\
    Std & 0.05  & 0.04  & 0.04  & 0.03  & 0.03  & 0.03  & 0.03  & 0.02  & 0.02 \\ \hline
    \multicolumn{10}{l}{\textbf{Model (B ) - With additive and ad-valorem transport costs} } \\ \hline
     \multicolumn{10}{l}{\textit{Multiplicative term} ($\widehat{\tau}^{adv}$)} \\ \hline
    Mean (in \%) & 5.42  & 3.08  & 4.02  & 3.31  & 2.79  & 2.49  & 2.68  & 1.95  & 2.22 \\
    Median (in \%) & 4.93  & 2.42  & 3.60  & 2.81  & 2.53  & 2.07  & 2.08  & 1.76  & 1.82 \\
    Std & 0.04  & 0.02  & 0.03  & 0.02  & 0.02  & 0.02  & 0.02  & 0.02  & 0.01 \\ \hline
       \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)}  \\ \hline
    Mean (in \%)  & 5.08  & 3.38  & 3.19  & 2.73  & 2.73  & 2.80  & 3.02  & 2.47  & 1.46 \\
    Median (in \%) & 2.94  & 2.27  & 2.06  & 1.70  & 1.82  & 2.19  & 2.16  & 1.89  & 0.76 \\
    Std & 0.09  & 0.05  & 0.04  & 0.04  & 0.04  & 0.04  & 0.04  & 0.03  & 0.02 \\ \hline
    \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
    Mean (in \%) & 14.51 & 10.05 & 10.47 & 14.62 & 8.37  & 8.02  & 8.41  & 6.40  & 5.23 \\
    Median (in \%)& 9.53  & 6.69  & 7.16  & 6.22  & 4.68  & 4.93  & 5.74  & 3.87  & 3.57 \\
    Std & 0.24  & 0.17  & 0.18  & 0.30  & 0.15  & 0.16  & 0.15  & 0.15  & 0.10 \\ \hline \hline
   \end{tabular}%
   }
\end{table}%

@@

To go further into characterizing the empirical relevance of the model with aditive costs only (Model (C)), we compare quality of fit diagnostic tests in this section. Table \ref{tab:3models_diagnosis_air} reports the results for air transport, and Table \ref{tab:3models_diagnosis_vessel} those for vessel transport. In both tables, we report the values of the $R^2$, the Standard Error of Regression (SER), the AIC criterion  and the log-likelihood (LL) value. We also report the value of the log-likelihood ratio that tests the quality of fit of the global model (Model (B)) compared to the other two models.

@@

    Model (B ) & \multicolumn{1}{c}{31387.3} & \multicolumn{1}{c}{35738.4} & \multicolumn{1}{c}{42535.8} & \multicolumn{1}{c}{52098.9} & \multicolumn{1}{c}{61343.7} & \multicolumn{1}{c}{74954.9} & \multicolumn{1}{c}{92758.6} & \multicolumn{1}{c}{95887.1} & \multicolumn{1}{c}{100155.4} \\

@@

    Model (B ) & \multicolumn{1}{c}{-15125.6} & \multicolumn{1}{c}{-17263.2} & \multicolumn{1}{c}{-20686.9} & \multicolumn{1}{c}{-25393.5} & \multicolumn{1}{c}{-30036.9} & \multicolumn{1}{c}{-36788.4} & \multicolumn{1}{c}{-45768.3} & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-49419.7} \\

@@

\begin{landscape}
\begin{table}[htbp]
  \centering
  \footnotesize{
  \caption{Quality-of-fit diagnostic tests of the three models (Vessel, 3-digit level)}
    \label{tab:3models_diagnosis_vessel}%
   \begin{tabular}{l|c c c c c c c c c}
    \hline \hline
    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
    \multicolumn{10}{l}{\textbf{$R^2$}} \\ \hline
    Model (A) & 0.450 & 0.415 & 0.427 & 0.456 & 0.438 & 0.401 & 0.378 & 0.350 & 0.339 \\
    Model (B) & 0.612 & 0.575 & 0.571 & 0.590 & 0.611 & 0.571 & 0.541 & 0.491 & 0.462 \\
    Model (C ) & 0.424 & 0.401 & 0.374 & 0.429 & 0.456 & 0.431 & 0.417 & 0.358 & 0.349 \\ \hline
    \multicolumn{10}{l}{\textbf{SER} } \\ \hline
    Model (A) & 0.576 & 0.620 & 0.569 & 0.592 & 0.615 & 0.652 & 0.673 & 0.740 & 0.758 \\
    Model (B) & 0.484 & 0.528 & 0.493 & 0.514 & 0.512 & 0.551 & 0.578 & 0.656 & 0.684 \\
    Model (C ) & 1.271 & 1.339 & 1.283 & 1.326 & 1.302 & 1.319 & 1.336 & 1.392 & 1.410 \\ \hline
    \multicolumn{10}{l}{\textbf{AIC criteria} } \\ \hline
    Model (A) & 33328.81 & 33010.27 & 40275.70 & 51142.62 & 60414.92 & 71365.89 & 85051.02 & 84789.89 & 88191.87 \\
    Model (B) & 27331.52 & 28067.31 & 34170.52 & 43664.74 & 49275.33 & 60475.91 & 73020.09 & 76161.33 & 80873.72 \\
    Model (C ) & 46082.40 & 44370.26 & 58829.71 & 71461.52 & 77052.41 & 88746.51 & 103310.93 & 101166.91 & 104290.27 \\ \hline
    \multicolumn{10}{l}{\textbf{Log-likelihood}}\\ \hline
    Model (A) & -16287.40 & -16129.13 & -19767.85 & -25169.31 & -29790.46 & -35263.95 & -42122.51 & -41998.95 & -43692.93 \\
    Model (B) & -12985.76 & -13353.65 & -16398.26 & -21171.37 & -23905.66 & -29490.96 & -35844.04 & -37418.66 & -39751.86 \\
    Model (C ) & -22674.20 & -21814.13 & -29045.86 & -35403.76 & -38125.20 & -43963.25 & -51245.46 & -50348.45 & -51783.14 \\ \hline
   \multicolumn{10}{l}{ \textbf{Test LL} } \\ \hline
    Stat LL ratio (B vs A) & 6603.28 & 5550.96 & 6739.18 & 7995.88 & 11769.59 & 11545.98 & 12556.94 & 9160.56 & 7882.15 \\
    $\#$ restrictions & 393   & 395   & 403   & 411   & 429   & 436   & 437   & 424   & 427 \\
    p-value & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00 \\ \hline
    Stat LL ratio (B vs C) & 19376,88 & 16920,95 & 25295,20 & 28464,79 & 28439,08 & 28944,59 & 30802,84 & 25859,58 & 24062,55 \\
    $\#$ restrictions & 393   & 395   & 403   & 411   & 429   & 436   & 437   & 424   & 427 \\
    p-value & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00 \\
\hline \hline
    \end{tabular}%
}
\end{table}%
\end{landscape}

The model with additive costs only (Model (C)) is consistently dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Model (A)), which is itself consistently dominated by the complete model (Model (B)), whatever the type of diagnostic test considered. That justify our choice to disregard Model (C) in the main text.

@@

\subsection{Transport Cost Estimates: Yearly Detailed Results}
In this section, we complement Table 1 of the main text by reporting the year-to-year results of the estimation driven at the 3-digit classification level. Table \ref{tab_oa:result_air_ally3} reports the results for each year over 1974-2013 for air transport; Table \ref{tab_oa:result_vessel_ally3} reports similar results for vessel. In both cases, we report the estimated values of the transport costs (weighed mean and median) when only ad-valorem costs are modeled (Model (A)) and when both additive and ad-valorem costs are modeled (Model (B)).

@@

The situation is is strikingly for primary goods only. In this case, it is in air transport that composition effects do matter (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (d), (e) and (f)). Furthermore, in air transport, composition effects matter by partially offsetting the decrease in the ``ceteris paribus'' transport costs (ie, implying a reduction in the ``raw'' transport costs over time much less pronounced than the fitted transport cost measures).

>>>>

    Standard Deviation & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline

@@

    Standard Deviation & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.01} \\ \hline

@@

    Standard Deviation & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.10} & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.06} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.05} \\  \hline \hline

@@

\begin{table}[htbp]
  \centering
   \footnotesize{
    \caption{Estimation results of the three models (Vessel, 3-digit level)}
  \label{tab:3models_estimation_results_vessel}%
    \begin{tabular}{l|c c c c c c c c c}
    \hline \hline
    \textbf{Year} & 1974  & 1980  & 1985  & 1990  & 1995  & 2000  & 2005  & 2010  & 2013 \\ \hline
    $\# $obs. & 19007 & 17356 & 23348 & 28383 & 32146 & 36090 & 41319 & 37748 & 38473 \\
    $\#$  origin countries & 154   & 163   & 171   & 179   & 201   & 206   & 206   & 198   & 203 \\
    $\#$  sectors & 239   & 232   & 232   & 232   & 228   & 230   & 231   & 226   & 224 \\ \hline
    \multicolumn{10}{l}{\textbf{Model (A) - Iceberg transport costs only} ($\widehat{\tau}^{ice}$)} \\ \hline
    Mean (in \%) & 9.79  & 6.53  & 6.88  & 5.67  & 5.14  & 5.10  & 5.47  & 3.99  & 3.60 \\
    Median (in \%)& 9.58  & 5.50  & 6.33  & 4.63  & 4.29  & 4.85  & 4.90  & 3.56  & 3.28 \\
    Standard Deviation & 0.05  & 0.04  & 0.04  & 0.03  & 0.03  & 0.03  & 0.03  & 0.02  & 0.02 \\ \hline
    \multicolumn{10}{l}{\textbf{Model (B ) - With additive and ad-valorem transport costs} } \\ \hline
     \multicolumn{10}{l}{\textit{Multiplicative term} ($\widehat{\tau}^{adv}$)} \\ \hline
    Mean (in \%) & 5.42  & 3.08  & 4.02  & 3.31  & 2.79  & 2.49  & 2.68  & 1.95  & 2.22 \\
    Median (in \%) & 4.93  & 2.42  & 3.60  & 2.81  & 2.53  & 2.07  & 2.08  & 1.76  & 1.82 \\
    Standard Deviation & 0.04  & 0.02  & 0.03  & 0.02  & 0.02  & 0.02  & 0.02  & 0.02  & 0.01 \\ \hline
       \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)}  \\ \hline
    Mean (in \%)  & 5.08  & 3.38  & 3.19  & 2.73  & 2.73  & 2.80  & 3.02  & 2.47  & 1.46 \\
    Median (in \%) & 2.94  & 2.27  & 2.06  & 1.70  & 1.82  & 2.19  & 2.16  & 1.89  & 0.76 \\
    Standard Deviation & 0.09  & 0.05  & 0.04  & 0.04  & 0.04  & 0.04  & 0.04  & 0.03  & 0.02 \\ \hline
    \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
    Mean (in \%) & 14.51 & 10.05 & 10.47 & 14.62 & 8.37  & 8.02  & 8.41  & 6.40  & 5.23 \\
    Median (in \%)& 9.53  & 6.69  & 7.16  & 6.22  & 4.68  & 4.93  & 5.74  & 3.87  & 3.57 \\
    Standard Deviation & 0.24  & 0.17  & 0.18  & 0.30  & 0.15  & 0.16  & 0.15  & 0.15  & 0.10 \\ \hline \hline
   \end{tabular}%
   }
\end{table}%

@@

To go further into the comparison of the empirical relevance of our three empirical models (A), (B) and (C), we compare quality of fit diagnostic tests in this section. Table \ref{tab:3models_diagnosis_air} reports the results for air transport, and Table \ref{tab:3models_diagnosis_vessel} those for vessel transport. In both tables, we report the values of the $R^2$, the Standard Error of Regression (SER), the AIC criterion  and the log-likelihood (LL) value. We also report the value of the log-likelihood ratio that tests the quality of fit of the global model (Model (B)) compared to the other two models.

@@

    Model (B) & \multicolumn{1}{c}{31387.3} & \multicolumn{1}{c}{35738.4} & \multicolumn{1}{c}{42535.8} & \multicolumn{1}{c}{52098.9} & \multicolumn{1}{c}{61343.7} & \multicolumn{1}{c}{74954.9} & \multicolumn{1}{c}{92758.6} & \multicolumn{1}{c}{95887.1} & \multicolumn{1}{c}{100155.4} \\

@@

    Model (B) & \multicolumn{1}{c}{-15125.6} & \multicolumn{1}{c}{-17263.2} & \multicolumn{1}{c}{-20686.9} & \multicolumn{1}{c}{-25393.5} & \multicolumn{1}{c}{-30036.9} & \multicolumn{1}{c}{-36788.4} & \multicolumn{1}{c}{-45768.3} & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-49419.7} \\

@@

\begin{landscape}
\begin{table}[htbp]
  \centering
  \footnotesize{
  \caption{Quality-of-fit diagnostic tests of the three models (Vessel, 3-digit level)}
    \label{tab:3models_diagnosis_vessel}%
   \begin{tabular}{l|c c c c c c c c c}
    \hline \hline
    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
    \multicolumn{10}{l}{\textbf{$R^2$}} \\ \hline
    Model (A) & 0.450 & 0.415 & 0.427 & 0.456 & 0.438 & 0.401 & 0.378 & 0.350 & 0.339 \\
    Model (B) & 0.612 & 0.575 & 0.571 & 0.590 & 0.611 & 0.571 & 0.541 & 0.491 & 0.462 \\
    Model (C ) & 0.424 & 0.401 & 0.374 & 0.429 & 0.456 & 0.431 & 0.417 & 0.358 & 0.349 \\ \hline
    \multicolumn{10}{l}{\textbf{SER} } \\ \hline
    Model (A) & 0.576 & 0.620 & 0.569 & 0.592 & 0.615 & 0.652 & 0.673 & 0.740 & 0.758 \\
    Model (B) & 0.484 & 0.528 & 0.493 & 0.514 & 0.512 & 0.551 & 0.578 & 0.656 & 0.684 \\
    Model (C) & 1.271 & 1.339 & 1.283 & 1.326 & 1.302 & 1.319 & 1.336 & 1.392 & 1.410 \\ \hline
    \multicolumn{10}{l}{\textbf{AIC criteria} } \\ \hline
    Model (A) & 33328.81 & 33010.27 & 40275.70 & 51142.62 & 60414.92 & 71365.89 & 85051.02 & 84789.89 & 88191.87 \\
    Model (B) & 27331.52 & 28067.31 & 34170.52 & 43664.74 & 49275.33 & 60475.91 & 73020.09 & 76161.33 & 80873.72 \\
    Model (C) & 46082.40 & 44370.26 & 58829.71 & 71461.52 & 77052.41 & 88746.51 & 103310.93 & 101166.91 & 104290.27 \\ \hline
    \multicolumn{10}{l}{\textbf{Log-likelihood}}\\ \hline
    Model (A) & -16287.40 & -16129.13 & -19767.85 & -25169.31 & -29790.46 & -35263.95 & -42122.51 & -41998.95 & -43692.93 \\
    Model (B) & -12985.76 & -13353.65 & -16398.26 & -21171.37 & -23905.66 & -29490.96 & -35844.04 & -37418.66 & -39751.86 \\
    Model (C) & -22674.20 & -21814.13 & -29045.86 & -35403.76 & -38125.20 & -43963.25 & -51245.46 & -50348.45 & -51783.14 \\ \hline
   \multicolumn{10}{l}{ \textbf{Test LL} } \\ \hline
    Stat LL ratio (B vs A) & 6603.28 & 5550.96 & 6739.18 & 7995.88 & 11769.59 & 11545.98 & 12556.94 & 9160.56 & 7882.15 \\
    $\#$ restrictions & 393   & 395   & 403   & 411   & 429   & 436   & 437   & 424   & 427 \\
    p-value & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00 \\ \hline
    Stat LL ratio (B vs C) & 19376,88 & 16920,95 & 25295,20 & 28464,79 & 28439,08 & 28944,59 & 30802,84 & 25859,58 & 24062,55 \\
    $\#$ restrictions & 393   & 395   & 403   & 411   & 429   & 436   & 437   & 424   & 427 \\
    p-value & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00 \\
\hline \hline
    \end{tabular}%
}
\end{table}%
\end{landscape}

The model with additive costs only (Model (C)) is consistently dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Model (A)), which is itself consistently dominated by the complete model (Model (B)), whatever the type of diagnostic test considered. That justifies our choice to disregard Model (C) in the main text.

@@

\subsection{Transport Cost Estimates: Yearly Detailed Results}
In this section, we complement Table 1 of the main text by reporting the year-to-year results of the estimation driven at the 3-digit classification level. Table \ref{tab_oa:result_air_ally3} reports the results for each year over 1974-2013 for air transport; Table \ref{tab_oa:result_vessel_ally3} reports similar results for vessel. In both cases, we report the estimated values of the transport costs (weighted mean and median) when only ad-valorem costs are modeled (Model (A)) and when both additive and ad-valorem costs are modeled (Model (B)).

@@

The situation is strikingly for primary goods only. In this case, it is in air transport that composition effects do matter (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (d), (e) and (f)). Furthermore, in air transport, composition effects matter by partially offsetting the decrease in the ``ceteris paribus'' transport costs (ie, implying a reduction in the ``raw'' transport costs over time much less pronounced than the fitted transport cost measures).
